---
title: "Stat-295-HW2"
author: "Sabahattin Alp Kocabaş, Başak Kabaloğlu, Efe Örencik"
date: "`r Sys.Date()`"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

-   **Loading necessary packages.**

    ```{r message=FALSE, warning=FALSE}
    library(ggplot2)
    library(tidyverse)
    ```

-   **Reading the data.**

    ```{r}
    data <- read.csv("social_network_ad.csv")
    ```

-   **Displaying the structure and summary of the data set.**

    ```{r}
    str(data)
    summary(data)
    head(data)
    ```

-   **Our observations for the dataset**

    The dataset has no missing values. The average age is approximately 37.66 years with a standard deviation of 10.48 years. Estimated salaries range from \$15,000 to \$150,000, with an average of \$69,742.50. About 35.75% of individuals made a purchase. Approximately 72.25% received a gift ticket.

-   **(i) Plot showing the relationship between Estimated Salary and Purchased.**

    ```{r message=FALSE}
    ggplot(data, aes(x =EstimatedSalary, y= Purchased)) +
      geom_point() +
      labs(title = "Relationship between Estimated Salary and Purchased",
           x = "Estimated Salary",
           y = "Purchased")+
      geom_smooth(method = "glm", se=FALSE, method.args = list(family = "binomial"))
    ```

-   **(i) Using geom_jitter() to make our graph more informative.**

    ```{r message=FALSE}
    ggplot(data, aes(x = EstimatedSalary, y= Purchased)) +
      geom_jitter(width = 0.5, height =0.08,alpha=0.6) + 
      geom_smooth(method = "glm", se=FALSE, method.args = list(family="binomial"))
    ```

-   **(ii) Fitting the logistic regression model.**

    ```{r}
    logit<- glm(Purchased ~ EstimatedSalary, data= data, family = binomial)
    summary(logit)
    ```

-   **(ii) Model Interpretation.**

    The logistic regression model estimates the relationship between EstimatedSalary and the probability of purchasing a product. The summary provides coefficients that can be interpreted as follows:

    Intercept (β0): The log-odds of purchasing when EstimatedSalary is zero

    EstimatedSalary (β1): The change in log-odds of purchasing for a one-unit increase in EstimatedSalary.

-   **(ii) Odds Ratio and Our Comment.**

    ```{r}
    odds_ratio <- exp(coef(logit))
    odds_ratio
    ```

    The odds of purchasing increase with increasing estimated salary.

-   **(iii) Calculating predicted probability with an estimated salary of \$22,000.**

    ```{r}
    salary <- data.frame(EstimatedSalary = 22000)
    predicted_prob <- predict(logit, salary, type = "response")
    predicted_prob
    ```

    The predicted probability of purchasing a product for an individual with an estimated salary of \$22,000 is approximately `r predicted_prob`.

-   **(iv) Converting the Gender column into binary variables (0 for female and 1 for male), and model of Gender.**

    ```{r}
    data$GenderBinary <- ifelse(data$Gender == "Male", 1, 0)

    logit_gender <- glm(Purchased ~ GenderBinary, family = binomial,data = data)
    summary(logit_gender)
    ```

-   **(iv) Odds Ratio.**

    ```{r}
    odds_ratios <- exp(coef(logit_gender))
    odds_ratios
    ```

    The odds ratio tells us how the odds of purchasing change for males compared to females. Since the odds ratio is less than 1, it suggests that males have a lower probability of purchasing compared to females.

-   **(v) New model to assess the impact of having a gift ticket on the likelihood of purchasing a product.**

    ```{r}
    logit_gift <- glm(Purchased ~  GiftTicket, data = data, family = binomial)

    summary(logit_gift)
    ```

-   **(v) Odds Ratio for GiftTicket.**

    ```{r}
    odds_ratio_gift <- exp(coef(logit_gift))
    odds_ratio_gift
    ```

    Since the odds ratio is greater than 1, it suggests that having a gift ticket increases the odds of purchasing.

-   **(vi) Computing the predicted probability of purchasing a product for each level of the GiftTicket variable.**

    ```{r}
    predictiondata <- data.frame(GiftTicket = c(0, 1))

    predictiondata$predicted_prob <- predict(logit_gift, newdata = predictiondata, type = "response")

    print(predictiondata)

    ```

# Question 2

-   **Our Plan.**

    For this question we chose the given airbnb dataset. We wanted our application to have three tabs,

    1-Summary Statistics,

    2-Interactive NYC Map,

    3-Filtered List.

    We made a brief comment for Summary statistics, and created a filtering system for Neighborhood Group, Room Type, Price Range variables for the interactive NYC Map. In addition to the filter containing the same variables for the Filtered List tab, we added another filter where we can select the columns we want to see in the list suitable for filtering.

<!-- -->

-   **Loading necessary packages.**

    ```{r message=FALSE}
    library(shiny)
    library(leaflet)
    library(dplyr)
    ```

-   **Load the data set.**

    ```{r}
    data <- read.csv("AB_NYC_2019.csv")
    ```

-   **Defining UI.**

    ```{r message=FALSE, warning=FALSE}
    ui <- navbarPage("Airbnb Listings for NYC",
      
      # Summary Statistics Tab
      tabPanel("Summary Statistics",
               fluidPage(titlePanel("Summary Statistics"),
                 mainPanel(verbatimTextOutput("summary"),
                   h3("Comments:"),
                   print("This table contains summary statistics for
                         various features of Airbnb listings. The dataset 
                         includes a total of 48,895 records. Prices vary widely,
                         ranging from a minimum of 0 to a maximum of 10,000 USD.
                         The average price is 152.7 USD, but the median price is 
                         106 USD,indicating a right-skewed distribution.")))),
      
      # Interactive NYC Map Tab
      tabPanel("Interactive NYC Map",
               fluidPage(titlePanel("NYC Map"),
                 sidebarLayout(sidebarPanel(selectInput("neighborhood_group", "Neighborhood Group:", 
                                 choices = unique(data$neighbourhood_group), 
                                 selected = "Manhattan"),
                     selectInput("room_type", "Room Type:", 
                                 choices = unique(data$room_type), 
                                 selected = "Entire home/apt"),
                     sliderInput("price", "Price Range:", 
                                 min = min(data$price), max = max(data$price), 
                                 value = c(min(data$price), max(data$price)))),
                   mainPanel(leafletOutput("map"))))),
      
      # Filtered List Tab
      tabPanel("Filtered List",
               fluidPage(titlePanel("Filtered List"),
                 sidebarLayout(sidebarPanel(selectInput("neighborhood_group_list", "Neighborhood Group:", 
                                 choices = unique(data$neighbourhood_group), 
                                 selected = "Manhattan"),
                     selectInput("room_type_list", "Room Type:", 
                                 choices = unique(data$room_type), 
                                 selected = "Entire home/apt"),
                     sliderInput("price_list", "Price Range:", 
                                 min = min(data$price), max = max(data$price), 
                                 value = c(min(data$price), max(data$price))),
                     checkboxGroupInput("columns", "Select Columns to Display:", 
                                        choices = names(data), 
                                        selected = names(data))),
                   mainPanel(tableOutput("filtered_list"))))))

    ```

-   **Defining server logic.**

    ```{r message=FALSE, warning=FALSE}
    server <- function(input, output, session) {
      
      # Summary Statistics of data 
      output$summary <- renderPrint({summary(data)})
      
      # Interactive NYC Map
      filtered_data <- reactive({data %>%
          filter(neighbourhood_group == input$neighborhood_group,
                 room_type == input$room_type,
                 price >= input$price[1],
                 price <= input$price[2])})
      
      output$map <- renderLeaflet({leaflet(filtered_data()) %>%
          addTiles() %>%
          addCircleMarkers(~longitude, ~latitude, 
                           popup = ~paste(name, "<br>", "Price: $", price, "<br>", "Room Type: ", room_type),
                           radius = 3, color = "blue", stroke = FALSE, fillOpacity = 0.7)})
      
      # Filtered List
      filtered_list_data <- reactive({data %>%
          filter(neighbourhood_group == input$neighborhood_group_list,
                 room_type == input$room_type_list,
                 price >= input$price_list[1],
                 price <= input$price_list[2]) %>%
          select(all_of(input$columns))})
      
      output$filtered_list <- renderTable({
        filtered_list_data()}, rownames = TRUE)}
    ```

-   **Running the app.**

    ```{r}
    shinyApp(ui = ui, server = server)
    ```
